Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images

### 2021-11-09 tex/paper.wecs/wecs_tgrs__trans08nov21.tex
sufoco-part1 (09nov21)

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
 \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath
%\usepackage{amsaddr}		
%\usepackage{amsthm}		
%\usepackage{amsaddr}		
\usepackage{amsfonts}

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

\def\bSig\mathbf{\Sigma}
\newcommand{\VS}{V\&S}
\usepackage{url}
\usepackage{soul} % allows to pass a line in words
\newcommand{\Z}{\mathbb{Z}} % Set of integers
\newcommand{\re}{\mathbb{R}} % Set of real numbers
\newcommand{\sgn}{\mathrm{sign}} % Signal operador
\newcommand{\ci}{\mathrm{i}}
\def\sT{\mbox{\tiny$T$}}
\def\E{\mbox{{\rm E\,}}}
\def\cov{\mbox{{\rm cov\,}}}
\def\corr{\mbox{{\rm corr\,}}}
\def\var{\mbox{{\rm var\,}}}
\newcommand{\vgamma}{\pmb{\gamma}}
\newcommand{\vrho}{\pmb{\rho}}
\newcommand{\vbeta}{\pmb{\beta}}
\newcommand{\vxi}{\pmb{\xi}}
\newcommand{\vepsilon}{\pmb{\epsilon}}
\def\shalf{\mbox{{\tiny$\frac{1}{2}$}}}
\def\half{\mbox{{\small$\frac{1}{2}$}}}
%\newtheorem{theorem}{Theorem}{\bf}{\rm}  
\newtheorem{lemma}{Lemma}

@@

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Rodney~V.~Fonseca, Rog\'{e}rio~G.~Negri,
        Alu\'{i}sio~Pinheiro,
        and~Abdourrahmane~Atto,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
\thanks{This work was supported by FAPESP grants 2016/24469-6 and 2018/04654-9 and CNPq grants 309230/2017-9 and 310991/2020-0.}
\thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
\thanks{A.~Atto is with the LISTIC - Polytech Annecy-Chamb\'{e}ry, Universit\'{e} de Savoie, 74944 Annecy le Vieux Cedex, France (e-mail: Abdourrahmane.Atto@univ-savoie.fr).}}
% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~X, No.~X, Month~20XX}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised sparse procedure to detect spatio-temporal change points on multi-temporal SAR images or even on sequences of very high resolution images. The procedure is based on multiscale approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients. We present two complimentary multiscale measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images. We show WECS performance on simulated multi-temporal image data. We also apply the proposed method to a time series of 84 satellite images in the border region of Brazil and the French Guiana. The images were captured from November 08, 2015 to December 09 2017.
\end{abstract}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{W}{e} propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images. WECS is based upon correlation screening for energy apportionment on wavelet approximations. The spatial character of the change detection is attained on pixel level. The method is fast, scalable, linearly updatable, and the resulting measures are sparse.

@@

Besides different methodological paradigms, several areas of application receive special attention.  For instance, urban change detection applications via polarimetric SAR (POLSAR) images are discussed in \cite{ansari2020urban}. \cite{song2018multi} discusses land cover change detection in mountainous terrain via multi-temporal and multi-sensor remote sensing images. \cite{ru2021multi} studies multi-temporal scene classification and scene change detection. Deforestation change detection is discussed by \cite{barreto2016deforestation}.

Wavelet methods present many advantages for a plethora of applications \cite{vidakovic1999statistical} thanks to wavelet capabilities in capturing multiscale/multiresolution information. Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical} use different wavelet methods for change detection in satellite images.

@@

The rest of the text goes as follows. Section \ref{section_method} introduces the problem and presents the proposed method. We show WECS performance on simulated multi-temporal image data in Section \ref{section_validation}. In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.  Section \ref{section_discussion} concludes the paper with a discussion.

\section{Methodology}\label{section_method}

Let $\mathcal{I}(1),\ldots,\mathcal{I}(m)$ be a set of matrices representing the images of some region of interest. These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate. Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}(m)$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time. We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}(m)$. Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}. After application of this wavelet transform to $\mathcal{I}(m)$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX(m)$, a smooth version of $\mathcal{I}(m)$ with the same dimension. The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX(m)$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}(m)$. Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX(m)$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.

>>>>

\documentclass[journal]{IEEEtran}

\ifCLASSINFOpdf
 \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage{amsmath}
%\usepackage{amsaddr}		
%\usepackage{amsthm}		
%\usepackage{amsaddr}		
\usepackage{amsfonts}

%DEFS. RODNEY%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\bSig\mathbf{\Sigma}
\newcommand{\VS}{V\&S}
\usepackage{url}
\usepackage{soul} % allows to pass a line in words
\newcommand{\Z}{\mathbb{Z}} % Set of integers
\newcommand{\re}{\mathbb{R}} % Set of real numbers
\newcommand{\sgn}{\mathrm{sign}} % Signal operador
\newcommand{\ci}{\mathrm{i}}
\def\sT{\mbox{\tiny$T$}}
\def\E{\mbox{{\rm E\,}}}
\def\cov{\mbox{{\rm cov\,}}}
\def\corr{\mbox{{\rm corr\,}}}
\def\var{\mbox{{\rm var\,}}}
\newcommand{\vgamma}{\pmb{\gamma}}
\newcommand{\vrho}{\pmb{\rho}}
\newcommand{\vbeta}{\pmb{\beta}}
\newcommand{\vxi}{\pmb{\xi}}
\newcommand{\vepsilon}{\pmb{\epsilon}}
\def\shalf{\mbox{{\tiny$\frac{1}{2}$}}}
\def\half{\mbox{{\small$\frac{1}{2}$}}}
%\newtheorem{theorem}{Theorem}{\bf}{\rm}  
\newtheorem{lemma}{Lemma}

@@

\usepackage{hyperref}

%DEFS. ROGERIO%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{graphicx}

\usepackage{subfigure}
\usepackage{multirow}
\usepackage{color,colortbl}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{cite}
\usepackage{balance}

\usepackage{bm,bbm}
\usepackage{datetime2}
\usepackage[binary-units]{siunitx}

\usepackage{mathrsfs}

\usepackage{placeins}

\usepackage[abs]{overpic}
\usepackage{pict2e}
%\usepackage{float}
\usepackage{microtype}
\usepackage{pdfpages}

\makeatletter
\newcommand{\thickhline}{\noalign {\ifnum 0=`}\fi \hrule height 1pt \futurelet \reserved@a \@xhline}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% correct bad hyphenation here
%\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\title{Wavelet Spatio-Temporal Change Detection on multi-temporal SAR images}

\author{Rodney~V.~Fonseca, Rog\'{e}rio~G.~Negri,
        Alu\'{i}sio~Pinheiro,
        and~Abdourrahmane~Atto,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
\thanks{This work was supported by FAPESP (Grants 2016/24469-6, 2018/04654-9 and 2021/01305-6) and CNPq (Grants 309230/2017-9 and 310991/2020-0).}
\thanks{R.~Fonseca and A.~Pinheiro are with the Department of Statistics, University of Campinas, Campinas 13083-859, Brazil (e-mails: ra192588@dac.unicamp.br; pinheiro@ime.unicamp.br).}
\thanks{A.~Atto is with the LISTIC - Polytech Annecy-Chamb\'{e}ry, Universit\'{e} de Savoie, 74944 Annecy le Vieux Cedex, France (e-mail: Abdourrahmane.Atto@univ-savoie.fr).}
\thanks{R.~G.~Negri are with the Department of Environmental Engineering, São Paulo State University, São José dos Campos, Brazil (e-mail: rogerio.negri@unesp.br).}
}
% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~X, No.~X, Month~20XX}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
We introduce WECS (Wavelet Energies Correlation Screening), an unsupervised sparse procedure to detect spatio-temporal change points on multi-temporal SAR images or even on sequences of very high resolution images. The procedure is based on multiscale approximation for the multi-temporal images, wavelet energy apportionment, and ultra-high dimensional correlation screening for the wavelet coefficients. We present two complimentary multiscale measures in order to detect sudden and/or cumulative changes, as well as for the case of stationary or non-stationary multi-temporal images. We show WECS performance on synthetic multi-temporal image data. We also apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana. The images were captured from November 08, 2015 to December 09 2017. \textcolor{magenta}{(lapidaremos ao fim)}
\end{abstract}

@@

\section{Introduction}

%%%RGN: Iniciar com uma visão geral do problema (detecćão de mudanća em dados multitemporais). O que é (a ideia)? Porque é importante? Mencionar abordagens supervisionada e não supervisionadaA inclusão de um/dois parágrafos seria suficiente. laplace

\textcolor{red}{\IEEEPARstart{C}{hange} detection... (completar)...}

@@

%%%RGN: Eu acho muito interessante que esta pesquisa esteja destinada à dados SAR (lida bem com dados ruidosos, imagens de uma única banda/polarizaćão e contempla uma aplicaćão real com dados Sentinel), no entanto, sinto falta de uma justificativa que relacione o método à dados SAR.

%%%RGN: Podemos omitir o "via polarimetric SAR (POLSAR)"?
Besides different methodological paradigms, several areas of application receive special attention.  For instance, urban change detection applications via polarimetric SAR (POLSAR) images are discussed in \cite{ansari2020urban}. \cite{song2018multi} discusses land cover change detection in mountainous terrain via multi-temporal and multi-sensor remote sensing images. \cite{ru2021multi} studies multi-temporal scene classification and scene change detection. Deforestation change detection is discussed by \cite{barreto2016deforestation}.

%%%RGN: o "wavelet methods" aqui não está claro se é destinado à detecćão de mudanćas.
Wavelet methods present many advantages for a plethora of applications \cite{vidakovic1999statistical} thanks to wavelet capabilities in capturing multiscale/multiresolution information. Their computational efficiency and sparseness are specially relevant for large images and other high-dimensional data \cite{morettin2017wavelets}. \cite{atto2012multidate}, \cite{bouhlel2015multivariate}, \cite{celik2009multiscale}, \cite{cui2012statistical} use different wavelet methods for change detection in satellite images.

%%%RGN: acrescentar algo sobre Screening? (dai ficam amarrados os conceitos de wavelets e screening, abrindo caminho para o wecs)

%%%RGN: acredito que esse paragrafo fica melhor aqui
%%%RGN: o acrônimo WECS só foi definido no abstract (geralmente é necessário definir no abstract e depois no corpo do texto com maior detalhe)
We propose here a novel method for unsupervised spatio-temporal change detection in multi-temporal SAR images. WECS is based upon correlation screening for energy apportionment on wavelet approximations. The spatial character of the change detection is attained on pixel level. The method is fast, scalable, linearly updatable, and the resulting measures are sparse.

@@

%%%RGN: Deixar para o final, pois a estrutura do paper pode ser alterada até a submissão.
\textcolor{magenta}{The rest of the text goes as follows. Section \ref{section_method} introduces the problem and presents the proposed method. We show WECS performance on synthetic multi-temporal image data in Section \ref{section_validation}. In Section \ref{section_realdata} we apply the proposed method to a time series of 85 satellite images in the border region of Brazil and the French Guiana, for  images captured from November 08, 2015 to December 09 2017.  Section \ref{section_discussion} concludes the paper with a discussion.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Methodology}\label{section_method}
%%%RGN: Temporário, depois alteramos.
\section{Wavelet Energies Correlation Screening}\label{section_method}

%%%%%%%%%%%%%%%%%%%%%
\subsection{Overview}\label{methodOverview}

Figure~\ref{figOverview} provides an overview of the proposed method for multitemporal image change detection.

%%%RGN: Vamos fazer um diagrama geral explicando como funciona o método? (usar o draw.io) Com base no diagrama fazemos uma discussão em alto nível. A próxima sećão vem formalizando os elementos deste diagrama. Se fizer uma versão inicial deste diagrama, eu ajudo a polir.

%-----------------------
\begin{figure}[hbt]
\centering
\includegraphics[width=0.485\textwidth]{../../figs/coming-soon.png}
\caption{Proposed method overview.}\label{figOverview}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%
\subsection{Formalization}\label{methodFormalization}

Let $\mathcal{I}(1),\ldots,\mathcal{I}(m)$ be a set of matrices representing the images of some region of interest. These images may be relative to one SAR channel or a combination of channels; this will be specified when appropriate. Our goal is twofold: to find possible points in time where some relevant changes might have taken place at the region represented in $\mathcal{I}(m)$, $m=1,\ldots,n$, and to find which regions are closely associated to the observed changes along time. We shall address these tasks by analyzing the bidimensional stationary discrete wavelet decomposition of $\mathcal{I}(m)$.

Stationary wavelets (also known as non-decimated or redundant wavelets) is a traditional de-noising method that can be efficiently applied to two-dimensional signals such as images \cite{coifman1995translation,atto2012multidate,atto2016wavelet}. After application of this wavelet transform to $\mathcal{I}(m)$ at some appropriate resolution level $J\geq 1$, one of its by-products is a matrix of so called approximation wavelet coefficients $\vX(m)$, a smooth version of $\mathcal{I}(m)$ with the same dimension. The higher $J\in\{1,\ldots,\log_2(k)\}$ is, the smoother $\vX(m)$ gets, where $k$ is the minimum between the numbers of rows and columns of $\mathcal{I}(m)$. Many other aspects can be involved in wavelet analysis of images (e.g., different types of wavelet transform, choice of wavelet basis, thresholding of detail coefficients, etc.), but in the current work we focus on $\vX(m)$, which provides a simple way of performing wavelet smoothing and that gives interesting results; extensions based on further aspects of wavelet transforms are straightforward.

We can then consider further apportioning the total $\mathbb{L}_2$ energy of $\left\{\vX(m)\right\}$ as
\begin{equation} 
\sum_{m=1}^n\|\vX(m)\|^2 = n\|\bar{\mathcal{I}}\|^2+2n\langle\bar{\vX}-\bar{\mathcal{I}},\bar{\mathcal{I}}\rangle+\sum_{m=1}^n\|\mathcal \vX(m)-\bar{\mathcal{I}}\|^2,\label{E:eq_ANOVAlogimage}
\end{equation}
where $\bar{\mathcal{I}}=n^{-1}\sum_{m=1}^n\mathcal{I}(m)$, $\bar{\vX}=n^{-1}\sum_{m=1}^n\vX(m)$ and $\|\vA\|^2=\sum_{i=1}^n\sum_{j=1}^m a_{ij}^2$ for a $n\times m$ matrix $\vA=[a_{ij}]$.

The last term in (\ref{E:eq_ANOVAlogimage}) measures deviations in time of $\vX(m)$ compared to an {\it average image} $\bar{\mathcal{I}}$, what motivates us to such deviations to detect change points in time. Since each element (pixel) in $\vX(m)$ also has a corresponding sequence of deviations in time, the relation of each local deviation to the overall measure can also be quantified, what allows us to detect changes in space. Such relation shall be computed in the present work with the Pearson correlation, what shares connections with the idea of feature screening employed in high-dimensional regression, as explained further. Other measures of change in time could be evaluated as well, such as squared differences of consecutive times $\|\mathcal \vX(m)-\vX(m-1)\|^2$, $d(\vX(m),\bar{\vX})$ with $d(\cdot,\cdot)$ denoting some distance measure (e.g., Hellinger, Kullback-Leibler, etc.), among many other possibilities.
